Personality and political beliefs across the lifespan

### 2019-12-11 scripts/manuscript.Rmd
added abstract

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Here is where we will write our abstract.

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

@@

To see whether there were significant relations between social conservatism and the Big Five traits, we then ran a multiple regression predicting social conservatism from the Big Five personality traits. We found that together the personality traits explain about 13.5% of the variance in social conservatism scores, *F*(5, 125) = 5.05, *p* < .001, with an adjusted $R^2$ of .14 (see Table 3). We found that agreeablessness, $\beta$ = 3.93, *t*(125) = 2.40, *p* = .018, and emotional stability, $\beta$ = 4.30, *t*(125) = 2.37, *p* = .020, significantly predict social conservatism, such that, controlling for the other personality traits, those who scored higher on agreeableness and emotional stability measures tended to also score higher on the social conservatism measure. Additionally, openness to experiences margionally predicted social conservatism, $\beta$ = -3.73, *t*(125) = -1.96, *p* = .052, such that, controlling for other personality traits, those who scored higher on conservatism tended to score lower on the openness to experience measure.

>>>>

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Since the development of the Five Factor Model as a framework for measuring and classifying personality, psychology research has been interested in using this framework to examine the relationship between personality and cognitions, mindsets, and behaviors. Prior research has attempted to predict political beliefs and affiliations from personality traits, finding that the personality trait of openness to experience is positively correlated with liberalness, while the trait of conscientiousness is positively correlated with conservativeness. The current study attempts to replicate these findings in a U.S. Amazon Mechanical Turk population using the Ten Item Personality Inventory to measure the Big Five personality traits (openness to experiences, conscientiousness, extraversion, agreeableness, and emotional stability) and a 12-item Social and Economic Conservatism Scale to measure the social and economic dimensions of conservatism. Consistent with past research, the current study found that openness to experience was a significant negative predictor of both economic and social conservatism. Contrary to past findings, however, we found that agreeableness and emotional stability were both significant positive predictors of social conservatism, and conscientiousness was not a significant predictor of either social or economic conservatism. Neither social nor economic extraversion were strongly correlated with any of the Big Five personality traits, but they were strongly positively correlated with each other.

keywords          : "keywords: personality, openness to experiences, conscientiousness, extraversion, agreeableness, emotional stability, conservatism, social conservatism, economic conservatism"
wordcount         : "208"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

@@

To see whether there were significant relations between social conservatism and the Big Five traits, we then ran a multiple regression predicting social conservatism from the Big Five personality traits. We found that together the personality traits explain about 13.5% of the variance in social conservatism scores, *F*(5, 125) = 5.05, *p* < .001, with an adjusted $R^2$ of .14 (see Table 3). We found that agreeablessness, $\beta$ = 3.93, *t*(125) = 2.40, *p* = .018, and emotional stability, $\beta$ = 4.30, *t*(125) = 2.37, *p* = .020, significantly predict social conservatism, such that, controlling for the other personality traits, those who scored higher on agreeableness and emotional stability measures tended to also score higher on the social conservatism measure. Additionally, openness to experiences marginally predicted social conservatism, $\beta$ = -3.73, *t*(125) = -1.96, *p* = .052, such that, controlling for other personality traits, those who scored higher on conservatism tended to score lower on the openness to experience measure.
